### 2017-05-24 paper/AlignedIsotropicGW.tex
Close #33, shared P(a).

% !TEX TS-program = pdflatexmk

% arXiver will snag the right plots so they show up on Twitter :)
%@arxiver{chi-eff-distributions.pdf,Wills_evidence_ratio_figure_with_mixture_models.png,six-way-O2-model-selection.pdf}

\documentclass{nature}

\usepackage{acronym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{color}
\usepackage{hyperref}
\usepackage{graphicx}

\newcommand{\chieff}{\chi_\mathrm{eff}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\diff}[2]{\frac{\dd #1}{\dd #2}}
\newcommand{\fixme}[1]{\textcolor{red}{FIXME: #1}}
\newcommand{\pali}{p_\mathrm{ali}}
\newcommand{\piso}{p_\mathrm{iso}}

\newcommand{\checkme}[1]{\textcolor{red}{#1}}
%\newcommand{\checkme}[1]{ #1 }

\newcommand{\OOneSigmaIsoAligned}{1.7}
\newcommand{\OOneOddsIsoAligned}{0.087}
\newcommand{\OTwoSigmaIsoAlignedMin}{2.9}
\newcommand{\OTwoOddsIsoAlignedMin}{0.0035}

\newcommand{\ilya}[1]{\textcolor{purple}{#1}}
\newcommand{\will}[1]{\textcolor{cyan}{#1}}

\newcommand{\plotone}[1]{\includegraphics[width=\columnwidth]{#1}}

% Journal Abbreviations
\newcommand{\aap}{Astron. and Astrophys.}
\newcommand{\apj}{Astrophys. J.}
\newcommand{\apjl}{Astrophys. J. Lett.}
\newcommand{\apjs}{Astrophys. J. Supp.}
\newcommand{\mnras}{Mon. Not. R. Astron. Soc.}
\newcommand{\nat}{Nat.}
\newcommand{\physrep}{Phys. Rep.}
\newcommand{\prd}{Phys. Rev. D}

\begin{document}

\acrodef{BBH}{binary black hole}
\acrodef{BH}{black hole}
\acrodef{EM}{electromagnetic}
\acrodef{GW}{gravitational wave}
\acrodef{O1}{first observing}
\acrodef{PE}{parameter estimation}

\title{Distinguishing Spin-Aligned and Isotropic Black Hole
  Populations With Gravitational Waves}

\author{Will M. Farr$^1$, Simon Stevenson$^1$, M. Coleman Miller$^2$,
  Ilya Mandel$^{1,3}$, Ben Farr$^4$ \& Alberto Vecchio$^1$}

\maketitle

\begin{affiliations}
\item Birmingham Institute for Gravitational Wave Astronomy and
  School of Physics and Astronomy, University of Birmingham,
  Birmingham, B15 2TT, United Kingdom
\item Department of Astronomy and Joint Space-Science
  Institute, University of Maryland, College Park, MD 20742--2421,
  United States
\item Kavli Institute for Theoretical Physics, Santa Barbara, CA 93106
\item Enrico Fermi Institute and Kavli Institute for Cosmological
  Physics, University of Chicago, Chicago, IL 60637, United States
\end{affiliations}

\begin{abstract}
  The first direct detections of gravitational
  waves\cite{2016PhRvL.116f1102A,2016PhRvL.116x1103A,O1-BBH} from
  merging binary black holes open a unique window into the binary
  black hole formation environment.  One promising signature of the
  formation environment is the angular distribution of the black hole
  spins; systems formed through dynamical interactions among
  already-compact objects are expected to have isotropic spin
  orientations\cite{SigurdssonHernquist:1993,1993Natur.364..421K,PZMcMillan:2000,Rodriguez:2015,Stone:2016,2016ApJ...832L...2R}
  whereas binaries formed from pairs of stars born together are more
  likely to have spins preferentially aligned with the binary orbit as
  a consequence of their joint evolution toward a binary black hole
  system\cite{TutukovYungelson:1973,TutukovYungelson:1993,Lipunov:1997,2016Natur.534..512B,Stevenson:2017,MandeldeMink:2016,Marchant:2016}.
  We consider existing gravitational wave measurements of the binary
  effective spin, the best-measured combination of spin
  parameters\cite{2016PhRvL.116x1102A,O1-BBH}, in the three likely
  binary black hole detections GW150914, LVT151012, and GW151226.  If
  binary black hole spin magnitudes extend to high values we show that
  the data already exhibit a $\OOneSigmaIsoAligned\sigma$
  ($\OOneOddsIsoAligned$ odds ratio\footnote{An odds ratio of $r$ with
    $r \ll 1$ is equivalent to $x \sigma$ with
    $x = \Phi^{-1}\left( 1 - r/2 \right)$, where $\Phi$ is the unit
    normal CDF.})  preference for an isotropic angular distribution.
  By considering the effect of 10 additional
  detections\cite{2016ApJ...833L...1A} of systems, we show that if all
  observations come from a single population such an augmented data
  set would enable at least a $\OTwoSigmaIsoAlignedMin\sigma$
  ($\OTwoOddsIsoAlignedMin$ odds ratio) distinction between the
  isotropic and aligned models for the assumed spin magnitude
  distributions, and in most cases better than $5\sigma$
  ($2.9 \times 10^{-7}$ odds ratio).  The existing preference for
  either an isotropic spin distribution or low spin magnitudes for the
  observed systems will be confirmed (or overturned) confidently in
  the near future by subsequent observations.
\end{abstract}

\acresetall{}

Following the detection of a merging \ac{BBH} system, \ac{PE}
tools\cite{2015PhRvD..91d2003V} compare model gravitational
waveforms\cite{2014PhRvD..89h4006P,2014PhRvD..89f1502T,2014PhRvL.113o1101H}
against the observed data to obtain a posterior distribution on the
parameters that describe the compact binary source.  The spin
parameter with the largest effect on waveforms, and a correspondingly
tight constraint from the data\cite{2016PhRvL.116x1102A}, is a
mass-weighted combination of the components of the dimensionless spin
vectors of the two black holes that are aligned with the orbital axis,
the ``effective spin,'' $-1 < \chieff < 1$ (see Methods Section
\ref{methsec:chieff-spin-magnitude}).

Figure \ref{fig:O1-posteriors} shows an approximation to the posterior
inferred on $\chieff$ for the three likely \ac{GW} detections
GW150914, GW151226 and LVT151012 from Advanced LIGO's first observing
run (O1)\cite{O1-BBH}.  Because samples drawn from the posterior on
$\chieff$ are not publicly released at this time, we have approximated
the posterior as a Gaussian distribution with the same mean and 90\%
credible interval, truncated to $-1 < \chieff < 1$.  None of the
$\chieff$ posteriors are consistent with two black holes with large
aligned spins, $\chi_{1,2} \gtrsim 0.5$; this contrasts with the large
spins inferred for the majority of black holes in X-ray binaries with
claimed spin measurements (see below).  The analysis here is
relatively insensitive to the precise details of the posterior
distributions; other conclusions are more sensitive.  In particular,
our Gaussian approximation does permit $\chieff = 0$ for GW151226
while the true posterior rules this out at high
confidence\cite{2016PhRvL.116x1103A,O1-BBH}.

\begin{figure}
  \plotone{../plots/chi-eff-mock-posteriors}
  \caption{\label{fig:O1-posteriors} \textbf{Approximate posteriors on
      $\chieff$ from the Advanced LIGO O1 run
      observations\cite{O1-BBH}.}  We approximate the posteriors
    reported using Gaussians with the same median and 90\% credible
    interval.  It is notable that none of the $\chieff$ posteriors
    support high \ac{BH} spin magnitudes with aligned spins, suggested
    by observations of stellar-mass black holes in X-ray
    binaries\cite{2015PhR...548....1M}.}
\end{figure}

Small values of $\chieff$ as exhibited in these systems can result
from either intrinsically small spins or larger spins whose direction
is mis-aligned with the orbital angular momentum of the binary (i.e.\
spin vectors with small $z$-components).  Mis-alignment is capable of
producing \emph{negative} values of $\chieff$, however, whereas
aligned spins will always have $\chieff \geq 0$.  This difference
provides strong discriminating power between the two angular
distributions, even without good information about the magnitude
distribution; to the extent that data favour negative $\chieff$ they
weigh heavily against aligned models.  To quantify the degree of
support for these two alternate explanations of small $\chieff$ values
in the merging \ac{BBH} population, we compared the Bayesian evidence
for various simple models of the spin population using the \ac{GW}
data set.

Each of our models for the merging \ac{BBH} spin population assumes
that the merging black holes are of equal mass (this is marginally
consistent with the three observations\cite{O1-BBH}, and the $\chieff$
distribution is not particularly sensitive to the mass ratio between
the merging objects---see Methods Section \ref{sec:mass-ratio}).  We
assume that the population spin distribution factorises into a
distribution for the spin magnitude $a$ and a distribution for the
spin angles.  Finally, we assume that the distribution of spins is
common to each component in a merging binary (in reality the
distributions of spin for each component in the binary may differ
systematically due to different formation and accretion histories, but
our analysis is not sensitive to the magnitude distribution at this
level; see below).  Choosing one of three magnitude distributions (see
Methods Section \ref{methsec:chieff-spin-magnitude}), ``low'' (mean
$a = 0.33$, standard deviation $0.24$), ``flat'' (mean $a = 0.5$,
standard deviation $0.29$), ``high'' (mean $a = 0.67$, standard
deviation $0.24$) and pairing with either an isotropic angular
distribution or a distribution that generates perfect alignment with
the positive $z$ axis yields six different models for the $\chieff$
distribution.  These models are shown in Figure
\ref{fig:chieff-distribution-models}.  While the Advanced LIGO
searches use spin-aligned templates they are efficient in detecting
misaligned binary black hole systems\cite{2016PhRvD..93l2003A}; we
assume here that the $\chieff$ distribution of observed sources
follows the true population.

\begin{figure}
  \plotone{../plots/chi-eff-distributions}
  \caption{\label{fig:chieff-distribution-models} \textbf{The models for the
    population distribution of $\chieff$ considered in this paper.}  In
    all models we assume that the binary mass ratio
    $q \equiv m_1/m_2 = 1$ and that the distribution of spin vectors
    is the same for each component.  The ``flat'' (blue lines),
    ``high,'' (green lines), and ``low'' (red lines) magnitude
    distributions are defined in Eq.\ \eqref{eq:magnitude-dists}.
    Solid lines give the $\chieff$ distribution under the assumption
    that the orientations of the spins are isotropic; dashed lines
    give the distribution under the assumption that both objects'
    spins are aligned with the orbital angular momentum.  The
    isotropic distributions are readily distinguished from the aligned
    distributions by the production of negative $\chieff$ values,
    while the distinction between the three models for the spin
    magnitude distribution is less sharp.}
\end{figure}

These magnitude distributions not meant to represent any particular
physical model, but rather to capture our uncertainty about the spin
magnitude distribution, discussed in detail below; neither
observations nor population synthesis codes can at this point
authoritatively suggest \emph{any} particular spin
distribution\cite{2015PhR...548....1M}.  Our models, however, allow us
to see how sensitive the $\chieff$ distribution is to spin alignment
given uncertainties about the spin magnitudes.

We fit hierarchical models of the three LIGO O1 observations using
these six different, zero-parameter population distributions (see
Section \ref{sec:hierarchical}).  We also fit three mixture models for
the population, where the spin magnitude distribution is fixed but the
angular distribution is a weighted sum of the isotropic and aligned
distribution.  The evidence, or marginal likelihood, for each of the
models is shown in Figure \ref{fig:O1-odds}.  For all three magnitude
distributions, the mixture models' posterior on the mixing fraction
peaks at 100\% isotropic, which explains why the zero-parameter,
pure-isotropic models are preferred over the single-parameter mixture
models for every magnitude distribution with this data set.  Not
surprisingly, given the small $\chieff$ values in the three detected
systems, the most-favoured model among those with an isotropic angular
distribution has the ``low'' magnitude distribution; the most favoured
model among those with an aligned distribution also has the ``low''
magnitude distribution.  The odds ratio between the ``low'' aligned
and ``low'' isotropic models is $\OOneOddsIsoAligned$, or
$\OOneSigmaIsoAligned\sigma$; thus the data favour isotropic spins
among our suite of models.  While the data favour spin amplitude
distributions with small spin magnitudes, note that a model with all
\ac{BBH} systems having zero spin is ruled out by the GW151226
measurements, which bound at least one black hole to have spin
magnitude $\geq 0.2$ at 99\% credibility\cite{2016PhRvL.116x1103A}.

\begin{figure}
  \plotone{../plots/Wills_evidence_ratio_figure_with_mixture_models}
  \caption{\textbf{Odds ratios among our models using the
      approximations to the posteriors on $\chieff$ from the O1
      observations shown in Figure \ref{fig:O1-posteriors}.}  The flat
    (``F''), high (``H''), and low (``L'') spin magnitude
    distributions (see Eq.\ \eqref{eq:magnitude-dists}) are paired
    with isotropic (``I'') and aligned (``A'') angular distributions,
    as well as a mixture model of the two (``$\mathrm{M}$'').  The
    most-favoured models have the ``low'' distribution of spin
    magnitudes.  The odds ratio between these models is
    $\OOneOddsIsoAligned$, or $\OOneSigmaIsoAligned\sigma$.  For all
    magnitude distributions the pure-isotropic models are preferred
    over the mixture models; correspondingly, the posterior on the
    mixture fraction peaks at 100\% isotropic.}
  \label{fig:O1-odds}
\end{figure}

Estimates of the rate of \ac{BBH} coalescences give a reasonable
chance of 10 additional \ac{BBH} detections in the next two
years\cite{O1-BBH,2016ApJ...833L...1A,2016ApJS..227...14A}.  Assuming
10 additional detections drawn from each of our six zero-parameter
models for the spin distribution in addition to the three existing
detections from O1, with observational uncertainties drawn randomly
from the three Gaussian widths used to approximate the $\chieff$
posteriors in Figure \ref{fig:O1-posteriors}\footnote{The measurement
  uncertainty in $\chieff$ depends on the other parameters of the
  merging \ac{BBH} system, particularly on the mass ratio.  Our
  assumption about future observational uncertainties is appropriate
  if the parameters of the three detected events are representative of
  the parameters of future detections. See Section
  \ref{sec:chi-eff-precision} for further discussion.}, we find the
odds ratios shown in Figure \ref{fig:O2-predictions}.  We find that
most scenarios with an additional 10 detections allow the simulated
angular distribution to be inferred with greater than $5\sigma$
($2.9 \times 10^{-7}$ odds) credibility; and in the most pessimistic
case the distinction is typically $\OTwoSigmaIsoAlignedMin\sigma$
($\OTwoOddsIsoAlignedMin$ odds ratio).  While such future detections
should permit a confident distinction between \emph{angular}
distributions, we would remain much less certain about the
\emph{magnitude} distribution among the three options considered here
until we have a larger number of
observations\cite{2017arXiv170306869F,2017arXiv170306223G}.  In Figure
\ref{fig:O2-predictions}, the odds ratio between different magnitude
distributions with the same angular distribution is much closer to
unity than the odds ratio between angular distributions.

\begin{figure}
  \plotone{../plots/six-way-O2-model-selection}
  \caption{\label{fig:O2-predictions} \textbf{Distribution of odds
      ratios predicted with 10 additional observations above the three
      discussed above.}  Each panel corresponds to additional
    observations drawn from one of the $\chieff$ distribution models.
    The model from which the additional observations are drawn is
    outlined in red.  The height of the blue bar gives the median odds
    ratio relative to the model from which the additional observations
    are drawn; the green line gives the 68\% ($1 \sigma$) symmetric
    interval of odds ratios over 1000 separate draws from the model
    distribution.  The closest median ratio between the most-favoured
    isotropic model and the most-favoured aligned model is
    $\OTwoOddsIsoAlignedMin$, corresponding to
    $\OTwoSigmaIsoAlignedMin\sigma$ preference for the correct angular
    distribution; most models result in more than $5\sigma$ preference
    for the correct angular distribution.  Because the three existing
    observations are included in each data set the ``correct'' model
    is not necessarily preferred over the others, particularly when
    that model uses the ``high'' magnitude distribution, which is
    strongly dis-favoured from the O1 observations alone.}
\end{figure}

Most of our resolving power for the spin angular distribution is a
result of the fact that our ``aligned'' models cannot produce
$\chieff < 0$ (see Figure \ref{fig:chieff-distribution-models}).  If
spins are intrinsically very small, with $a \lesssim 0.3$, then it is
no longer possible to resolve the negative effective spin with a small
number of observations.  As noted below, however, spins observed in
X-ray binaries are typically large.  Additionally, models which do not
permit \emph{some} spins with $\chieff \gtrsim 0.1$ are ruled out by
the GW151226 observations\cite{2016PhRvL.116x1103A}.  An ``aligned''
model with spin magnitudes from our ``flat'' distribution but
permitting spin vectors oriented anti-parallel to the orbital angular
momentum (leading to the possibility of positive \emph{or} negative
$\chieff$) can only be distinguised from an isotropic true population
at $\sim 3 \sigma$ after 10--20
observations\cite{2017CQGra..34cLT01V}; our flat aligned model can be
distinguished from such a population at better than $5\sigma$ (odds
$< 10^{-8}$) after 10 observations, emphasising the information
content of the bound $\chieff > 0$ for our aligned models.

Observational data on spin magnitudes in black hole systems is
sparse\cite{2015PhR...548....1M}.  Most of the systems studied are
low-mass X-ray binaries rather than the high-mass X-ray binaries that
are likely to be the progenitors of double black hole binaries.  In
addition, there are substantial systematic errors that can complicate
these analyses\cite{2015PhR...548....1M} and selection effects could
yield a biased distribution. Nonetheless, if we take the reported spin
magnitudes as representative then we find that there is a preference
for high spins; for example, 14 of the 19 systems with reported spins
have dimensionless spin parameters in excess of 0.5.  It is usually
argued that the masses and spin parameters of stellar-mass black holes
are unlikely to be altered significantly by
accretion\cite{1999MNRAS.305..654K}, but this may not be true for all
systems\cite{2003MNRAS.341..385P,2015ApJ...800...17F}.  Thus the
current spin parameters are probably are close to their values upon
core collapse, at least in high-mass X-ray binaries.  However, the
specific processes involved in the production of black hole binaries
from isolated binaries could alter the spin magnitude distribution of
those holes relative to the X-ray binary systems; for example, close
tidal interactions could spin up the core, or stripping of the
envelope could reduce the available angular
momentum\cite{2016MNRAS.462..844K,2017arXiv170200885Z,2017arXiv170203952H}.

The spin directions of binary black holes formed dynamically through
interactions in dense stellar
environments\cite{SigurdssonHernquist:1993,1993Natur.364..421K,PZMcMillan:2000,Rodriguez:2015,Stone:2016}
are expected to be isotropic given the absence of a preferred
direction\cite{2016ApJ...832L...2R} and the persistence of an
isotropic distribution through post-Newtonian
evolution\cite{2004PhRvD..70l4020S,2007ApJ...661L.147B}.

The spin directions in isolated
binaries\cite{TutukovYungelson:1973,TutukovYungelson:1993,Lipunov:1997,2016Natur.534..512B,Stevenson:2017,MandeldeMink:2016,Marchant:2016}
are usually expected to be preferentially aligned.  Despite observed
spin-orbit misalignments in both massive stellar
binaries\cite{Albrecht:2009,2014ApJ...785...83A} and BH X-ray
binaries\cite{Orosz:2001,Martin:2008b,Martin:2008,MorningstarMiller:2014},
mass accretion and tidal interactions will tend to realign the binary.
On the other hand, a supernova natal kick (if any) can change the
orbital plane and misalign the
binary\cite{2000ApJ...541..319K,2013PhRvD..87j4028G}; the supernova
can also tilt the spin angle, as in the double pulsar
J0737-3039\cite{2011ApJ...742...81F}; and evolutionary processes, such
as wind-driven mass loss and post-collapse fallback, can couple the
spin magnitude and direction distributions, contrary to our simplified
assumptions.  A small misalignment at wide separation can also evolve
to a more significant misalignment in component spins as the binary
spirals in through \ac{GW} emission\cite{2015PhRvD..92f4016G}, but
$\chieff$ is approximately conserved through this evolution.

\begin{addendum}
\item We thank Richard O'Shaughnessy, Christopher Berry, Davide
  Gerosa, and Salvatore Vitale for discussions and comments on this
  work.  WF, SS, IM and AV were supported in part by the STFC.  MCM
  acknowledges support of the University of Birmingham Institute for
  Advanced Study Distinguished Visiting Fellows program.  SS and IM
  acknowledge support from the National Science Foundation under Grant
  No. NSF PHY11-25915.
\item[Supplementary Information] Supplementary Information is linked
  to the online version of the paper at
  \url{http://www.nature.com/nature}.
\item[Author Contributions] All authors contributed to the work
  presented in this paper.
\item[Author Information] Reprints and permissions information is
  available at \url{http://www.nature.com/reprints}.  The authors
  declare no competing financial interests.  Correspondence and
  requests for materials should be addressed to
  \href{mailto:w.farr@bham.ac.uk}{w.farr@bham.ac.uk}.
\end{addendum}

\begin{methods}

% Reset the Figure Counter
\setcounter{figure}{0}
\renewcommand{\figurename}{Extended Data Figure}

\section{Code Availability}

This analysis used the Julia language\cite{Julia}, Python libraries
NumPy and SciPy\cite{NumPy,SciPy}, the plotting library
Matplotlib\cite{Matplotlib}, and performed computations in IPython
notebooks\cite{IPython}.  A repository containing the code and
notebooks used for this analysis, together with the \LaTeX{} source
for this document, can be found under an open-source ``MIT'' license
at \url{https://github.com/farr/AlignedVersusIsoSpin}.

\section{Effective Spin and Spin Magnitude Distributions}
\label{methsec:chieff-spin-magnitude}

The effective spin is defined by\cite{2016PhRvL.116x1102A} 
\begin{equation}
  \chieff = \frac{c}{GM} \left( \frac{\vec{S}_1}{m_1} + \frac{\vec{S}_2}{m_2}
  \right) \cdot \frac{\vec{L}}{\left| \vec{L} \right|} \equiv \frac{1}{M} \left( m_1 \chi_1 + m_2 \chi_2 \right),
\end{equation}
where $m_{1,2}$ are the gravitational masses of the more-massive (1)
and less-massive (2) components, $M = m_1 + m_2$ is the total mass,
$\vec{S}_{1,2}$ are the spin angular momentum vectors of the black
holes in the binary, $\vec{L}$ is the orbital angular momentum vector,
assumed to point in the $\hat{z}$ direction, and
$0 \leq \chi_{1,2} \leq 1$ are the corresponding dimensionless
projections of the individual \ac{BH} spins.  Because the
dimensionless spin parameter,
\begin{equation}
  \label{eq:a-def}
  a_{1,2} = \frac{c}{G m_{1,2}^2} \left|\vec{S}_{1,2} \right|,
\end{equation}
of each black hole is bounded by $0 \leq a_{1,2} < 1$, the projections
along the orbital axis are bounded by $-1 < \chi_{1,2} < 1$, and
$-1 < \chieff < 1$.

We form the population distributions of $\chieff$ shown in Figure
\ref{fig:chieff-distribution-models} by assuming that each black hole
in a binary has a dimensionless spin magnitude drawn from one of three
distributions,
\begin{equation}
  \label{eq:magnitude-dists}
  p(a) = \begin{cases}
    2\left(1-a \right) & \textnormal{``low''} \\
    1 & \textnormal{``flat''} \\
    2 a & \textnormal{``high''}
  \end{cases},
\end{equation}
referred to as ``low,'' ``flat,'' and ``high'' in the text above.

\section{Mixture model}

While we carried out Bayesian comparisons between isotropic and aligned spin distributions under various assumptions, a preference for one of the considered models over the others does not necessarily indicate that it is the correct model.  All of the considered models could be inaccurate for the actual distribution, especially since all of the considered models are based on a number of additional assumptions, such as decoupled spin magnitude and spin misalignment angle distributions and identical distributions for primary and secondary spins.

We now partly relax the simplified assumptions made earlier by considering the possibility that the true distribution of BBH spin-orbit misalignments observed by LIGO is a mixture of binaries with aligned spins and binaries with isotropic spins.

%
\begin{figure}
\centering
\includegraphics[width=\textwidth]{../plots/posterior_on_isotropic_fraction.png}
\caption{\textbf{Fraction of the BBH population coming from an isotropic
  distribution under a mixture model.} The dotted line shows the flat
  prior on the fraction of BBHs coming from an isotropic distribution,
  $f_i$, under the mixture model. The 3 red lines show the posterior
  on $f_i$ after O1 with our various assumptions regarding BH spin
  magnitudes.  The solid line shows the posterior assuming that all
  BHs have their spin magnitude drawn from the ``flat''
  distribution. The dashed line assumes the ``high'' BH spin magnitude
  distribution $p(a) = 2a$. The dot-dash line assumes the ``low''
  distribution $p(a) = 2(1-a)$.  We see that for a wide range of
  assumptions regarding BH spin magnitudes, the fraction coming from
  an isotropic distribution $f_i$ peaks at 1.}
\label{fig:mixture_fraction_posterior}
\end{figure}
%

%-- could also use the cumulative posterior
%%
%\begin{figure}
%\centering
%\includegraphics[width=0.45\textwidth]{../plots/isotropic_fraction_cumulative_posterior.png}
%\caption{\textbf{Fraction of the BBH population coming from an isotropic distribution} The dotted (blue) line shows the cumulative distribution for the flat prior on the fraction of BBHs coming from an isotropic distribution $f_i$. The solid (red) line shows the cumulative distribution for the posterior after O1, assuming that all BHs have their spin magnitude drawn from a uniform distribution. The dashed line assumes a `increasing' distribution $p(a) = 2a$ for BH spin magnitudes, whilst the dot-dash line assumes a `decreasing' distribution $p(a) = 2(1-a)$. We see that regardless of our assumption regarding BH spin magnitudes, there is a preference for a large fraction coming from an isotropic distribution.}
%\label{fig:mixture_fraction_posterior}
%\end{figure}
%

We also fit a mixture model\cite{Stevenson:2017spin} (labelled model
'M' in Figure~\ref{fig:O1-odds}) where a fraction $f_i$ of BBHs have
spins drawn from an isotropic distribution, whilst a fraction
$1 - f_i$ have their spins aligned with the orbital angular
momentum. We assume a flat prior on the fraction $f_i$. To test the
robustness of our result, we vary the distribution we assume for BH
spin magnitude distributions as with the aligned and isotropic
models. We use the ``flat'', ``high'' and ``low'' distributions
(Equation~\ref{eq:magnitude-dists}), assuming all BHs have their spin
magnitude drawn from the same distribution for both the aligned and
isotropic populations. We calculate and plot the posterior on $f_i$
given by Equation~\ref{eq:hierarchical-posterior} ($f_i=\lambda$ in
the derivation) in Figure~\ref{fig:mixture_fraction_posterior}. We
find the mean fraction of BBHs coming from an isotropic distribution
is 0.63, 0.71 and 0.78 assuming the ``low'', ``flat'' and ``high''
distributions for spin magnitudes respectively, compared to the prior
mean of 0.5. The lower 90\% limits are 0.26, 0.39 and 0.52
respectively, compared to the prior of 0.1. In all cases, the
posterior peaks at $f_i = 1$. Thus, for these spin magnitude
distributions we find that the current O1 LIGO observations constrain
the majority of BBHs to have their spins drawn from an isotropic
distribution. The evidence ratios of these models to the isotropic
distribution with ``flat'' spin magnitudes are 0.85, 0.39 and 0.19 for
the ``low'', ``flat'' and ``high'' spin magnitude models. Thus we
cannot rule out a mixture with the current data.

\section{Hierarchical Modelling} 
\label{sec:hierarchical}

LIGO measures $\chieff$ better than any other spin parameter, but
still with significant uncertainty, so we need to properly incorporate
measurement uncertainty in our analysis; thus our analysis must be
\emph{hierarchical}\cite{2010ApJ...725.2166H,2010PhRvD..81h4029M}.
In a hierarchical analysis, we assume that each event has a true, but
unknown, value of the effective spin, drawn from the population
distribution, which may have some parameters $\lambda$; then the
system is observed, represented by the likelihood function, which
results in a distribution for the true effective spin (and all other
parameters describing the system) consistent with the data.
Combining, the joint posterior on each system's $\chieff^i$ parameters
and the population parameters $\lambda$ implied by a set of
observations each with data $d^i$, is
\begin{equation}
  p\left( \left\{ \chieff^i \right\}, \lambda \mid \left\{ d^i \right\} \right) \propto \left[ \prod_{i=1}^{N_\mathrm{obs}} p\left(d^i \mid \chieff^i \right) p\left( \chieff^i \mid \lambda \right) \right] p\left(\lambda\right).
\end{equation}

The components of this formula are
\begin{itemize}
\item The GW (marginal) likelihood, $p\left(d \mid \chieff\right)$.
  Here we use ``marginal'' because we are (implicitly) integrating
  over all parameters of the signal but $\chieff$.  Note that it is
  the likelihood rather than the posterior that matters for the
  hierarchical analysis; if we are given posterior distributions or
  posterior samples, we need to re-weight to ``remove'' the prior and
  obtain the likelihood.
\item The population distribution for $\chieff$,
  $p\left( \chieff \mid \lambda \right)$.  This function can be
  parameterised by population-level parameters, $\lambda$.  (In the
  cases discussed above, there are no parameters for the population.)
\item The prior on the population-level parameters, $p(\lambda)$.
\end{itemize}
If we do not care about the individual event $\chieff$ parameters, we
can integrate them out, obtaining
\begin{equation}
  p\left( \lambda \mid \left\{ d^i \right\} \right) \propto \left[ \prod_{i=1}^{N_\mathrm{obs}} \int \dd \chieff^i \, p\left(d^i \mid \chieff^i \right) p\left( \chieff^i \mid \lambda \right) \right] p\left(\lambda\right).
\end{equation}
If we are given posterior samples of $\chieff^{ij}$ ($i$ labels the
event, $j$ labels the particular posterior sample) drawn from an
analysis using a prior $p\left( \chieff \right)$, then we can
approximate the integral by a re-weighted average of the population
distribution over the samples (here $p\left( \chieff^{ij} \right)$ is
the prior used to produce the posterior samples):
\begin{equation}
  p\left( \lambda \mid \left\{ d^i \right\} \right) \propto \left[ \prod_{i=1}^{N_\mathrm{obs}} \frac{1}{N_i} \sum_{j=1}^{N_i} \frac{p\left( \chieff^{ij} \mid \lambda \right)}{p\left( \chieff^{ij} \right)} \right] p\left(\lambda\right).
  \label{eq:hierarchical-posterior}
\end{equation}

\subsection{Order of Magnitude Calculation}
\label{sec:om-odds-ratio}

It is possible to estimate at an order-of-magnitude level the rate at
which evidence accumulates in favour of or against the isotropic
models as more systems are detected.  Based on Figure
\ref{fig:chieff-distribution-models}, approximate the isotropic
population $\chieff$ distribution as uniform on
$\chieff \in \left[ -0.25, 0.25 \right]$ and the aligned population
$\chieff$ distribution as uniform on
$\chieff \in \left[0, 0.5\right]$.  Then the odds ratio between the
isotropic and aligned models for each event is approximately
\begin{equation}
  \label{eq:approx-odds}
  \frac{p\left( d \mid I \right)}{p\left( d \mid A \right)} \simeq
  \frac{P\left( -0.25 \leq \chieff \leq 0.25 \right)}{P\left( 0 \leq \chieff \leq 0.5 \right) },
\end{equation}
where $P\left( A \leq \chieff \leq B \right)$ is the posterior
probability (here used to approximate the likelihood) that $\chieff$
is between $A$ and $B$.  Using our approximations to the $\chieff$
posteriors described above, this gives an odds ratio of $5$ in favour
of the isotropic models, which is about a factor of two smaller than
the ratio in the more careful calculation described above.  This is a
satisfactory answer at an order-of-magnitude level.

If the true distribution is isotropic and follows this simple model,
and our measurement uncertainties on $\chieff$ are $\simeq 0.1$, then
the geometric mean of each subsequent measurement's contribution to
the overall odds is $\sim 3$.  After ten additional events, then, the
odds ratio becomes $5 \times 3^{10} \simeq 3 \times 10^{5}$, or
$4.6 \sigma$, consistent with the results of the more detailed
calculation described above.  If the true distribution of spins
becomes half as wide ($\chieff \in [-0.125, 0.125]$ for isotropic and
$\chieff \in [0, 0.25]$ for aligned spins), with the same
uncertainties, then the existing odds ratio becomes $1.08$, and each
subsequent event drawn from the isotropic distribution contributes on
average a factor of $1.6$.  In this case, after 10 additional events,
the odds ratio becomes $150$, or $2.7\sigma$.  With small spin
magnitudes, our angular resolving power vanishes, as discussed in more
detail in Section \ref{sec:smallspins}.

\section{Effect of small spin magnitudes}
\label{sec:smallspins}

In the main text we considered three models for BH spin magnitudes:
``low'', ``flat'' and ``high''. These were intended to capture some of
the uncertainty regarding the BH spin magnitude distribution.

Here we extend the ``low'' model as:
%
\begin{equation}
p(a) \propto (1 - a)^{\alpha}
\label{eq:lowspinalpha}
\end{equation}
%

When $\alpha = 0$, this recovers the ``flat'' distribution, whilst
$\alpha = 1$ recovers the ``low'' distribution. For higher values of
$\alpha$, this distribution becomes more peaked towards
$a = 0$.

%
\begin{figure}
\centering
\includegraphics[width=\textwidth]{../plots/sigma_v_alpha.png}
\caption{\textbf{Effect of small spins on evidence ratio of aligned
    against isotropic models.} The blue line shows the evidence ratio
  (plotted as the equivalent sigma) between a model where all systems
  are aligned, versus one where all systems are from an isotropic
  distribution as a function of the power law $\alpha$ corresponding
  to Equation~\ref{eq:lowspinalpha}.  The top axis shows the mean spin
  magnitude $\bar{a}$ which this $\alpha$ corresponds to. We see that
  for mean spin magnitudes $\lesssim 0.2$ we find no evidence for
  either distribution over the other.}
\label{fig:smallspinsalpha}
\end{figure}
%

In Figure~\ref{fig:smallspinsalpha} we plot the evidence ratio of
aligned to isotropic distributions (plotted as the equivalent sigma)
with spin magnitudes given by this model with $\alpha$ in the range
$0$ -- $6$. The top axis shows the mean spin magnitude that value of
$\alpha$ corresponds to (e.g., for the ``flat'' distribution
$\alpha = 0$, the mean spin magnitude is 0.5). We see that if typical
BH spins are $\lesssim 0.2$ we have no evidence for one model over the
other.

\section{Mass Ratio}
\label{sec:mass-ratio}

Figure \ref{fig:mass-ratio-sensitivity} shows the distributions of
$\chieff$ that would obtain with a mass ratio $q = m_2/m_1 = 0.5$
compared to the distributions with $q = 1$ used above.  The details of
the distribution are sensitive to the mass ratio, but in our analysis
we are primarily sensitive to the changing \emph{sign} of $\chieff$
under the isotropic models.  This latter property is insensitive to
mass ratio.  As an example, the distinction between the three
different spin amplitude distributions after ten additional detections
is quite weak compared to the aligned/isotropic distinction in Figure
\ref{fig:O2-predictions}.  The differences in the $\chieff$
distribution between $q = 1$ and $q = 0.5$ are even smaller than the
differences between the different magnitude distributions.

\begin{figure}
  \plotone{../plots/chi-eff-distributions-q05}
  \caption{\textbf{Distributions of $\chieff$ assuming all merging
      black holes have equal masses ($q=1$) or a 2:1 mass ratio
      ($q = 0.5$).}  The details of the distribution are sensitive to
    the mass ratio, but in our analysis we are primarily sensitive to
    the changing \emph{sign} of $\chieff$ under the isotropic models.
    This latter property is unchanged under changing mass ratio.}
  \label{fig:mass-ratio-sensitivity}
\end{figure}

\section{Approximations in the Gravitational Waveform}
The model waveforms used to infer the $\chieff$ of the three LIGO
events incorporate approximations to the true behaviour of the merging
systems that are expected to break down for sufficiently high
mis-aligned spins.  The effect of these approximations on inference on
the parameters describing GW150914 has been investigated in
detail\cite{2016arXiv161107531T}.  For this source, statistical
uncertainties dominate over any waveform systematics.  Detailed
comparisons with numerical relativity computations using no
approximations to the dynamics\cite{2016PhRvD..94f4035A} also suggest
that statistical uncertainties dominate the systematics for this
system.  Systematics may dominate for signals with this large SNR
($\simeq 23$) when the source is edge-on or has high
spins\cite{2016arXiv161107531T}.  The other two events discussed in
this paper are at much lower SNR, with correspondingly larger
statistical uncertainties, and are probably similarly oriented and
with similarly small spins, so we do not expect systematic
uncertainties to dominate.

We assume here that measurements made in the future are not dominated
by systematic errors, but this assumption would need to be revisited
for high-SNR, edge-on, or high-spin sources detected in the future.

\section{Precision of $\chieff$ measurements}
\label{sec:chi-eff-precision}
Throughout this work we have made the simplifying assumption that the
precision to which $\chieff$ can be constrained for individual binaries
is independent of the binary's properties.  In practice, our ability to
constrain $\chieff$ \emph{is} dependent on the system's properties, in
particular its true $\chieff$ and mass ratio, which we illustrate in
figure \ref{fig:chi-eff-constraints}.

For this figure a detected population\footnote{We qualify a system as
  ``detected'' if it produces a SNR above $8$ in the second-loudest
  detector to select only coincident events.} of 500 binaries was
selected from a population with component masses distributed uniformly
between $1$ and $30~\mathrm{M}_\odot$ with
$m_1 + m_2 < 30~\mathrm{M}_\odot$, locations distributed uniformly in
volume, and orientations distributed isotropically.  Data were
simulated for each binary, and posteriors were estimated using the
LIGO-Virgo parameter estimation library
\texttt{LALInference}\cite{2015PhRvD..91d2003V} using inspiral-only
waveform models (merger and ringdown effects can provide additional
information for some binaries, but we ignore those effects here).
$\chieff$ is better constrained for binaries with low (negative)
effective spins and low mass ratios.

By neglecting these effects, We do not expect to qualitatively affect
out conclusions, though they could affect predictions for the total
number of detections necessary to constrain the population.  For
example, if the universe preferentially forms asymmetric binaries with
low mass ratios, individual $\chieff$ constraints will be systematically
better, reducing the number of binaries needed to infer the properties
of the population.

\begin{figure}
  \plotone{../plots/chi-eff-90cl}
  \caption{\textbf{Widths of the 90\% credible intervals for $\chieff$
      for 500 binaries in a simulated detected population.} $\chieff$
    is better constrained for systems with low $\chieff$ and low mass
    ratio.}
  \label{fig:chi-eff-constraints}
\end{figure}

\end{methods}

\bibliographystyle{naturemag}
\bibliography{AlignedIsotropicGW}

\end{document}